
\let\textcircled=\pgftextcircled
\chapter{System design}

\initial{A}s mentioned in Section \ref{sec:methodology}, the design of the Obstacle Collision Avoidance System will follow the Systems Engineering approach.
The main reason is that Systems Engineering provides some methods that prevent the errors with the highest consequences when the system to be designed is complex.
As explained by Rolls-Royce Global Chief of Systems Engineering \cite{beasley2015}:
\begin{quote}
	\itshape
	Systems Engineering collects and organises all the information needed to understand the whole problem, explores it from all angles, and then finds the most appropriate system solution.
\end{quote}

Furthermore, A key study published through INCOSE \cite{incoseuk2016} looked at the phase of detection of errors, and the consequent cost of fixing them.
Cost modelling was validated against a cross-industry range of defence and aerospace projects.
Figure \ref{fig:incose} shows the results of the study.

\input{./figures/incose.tex}

Hence, in the present chapter, some of the most relevant Systems Engineering tools from the NASA Systems Engineering Handbook \cite{nationalaeronauticsandspaceadministration2007} will be applied


\section{Requirements capture} \label{sec:sysReqs}

The design process for a system is requirement driven, since the requirements are what will define the cost, design, schedule\ldots\ % Space after \ldots
A requirement is a statement about or a characteristic of something that is needed.

Requirements can be derived from a variety of sources, like customer needs, stakeholders, regulations, procedures, constrains, etc.
However, for this project, customers and stakeholders will be disregarded (since none exist) and the motivation as stated in Section \ref{sec:motivation} will be used instead.

In the present section some requirements will be posed, but only those that directly apply to the OCAS subsystem or its interfaces, since the platform is considered to be completely functional prior to the introduction of the solution (following the modularity concept).

\input{./tables/sysReqs.tex}

Notice that Table \ref{tab:sysReqs} is not static, and should be updated during the design process, since some of the tools of Systems Engineering are designed to expose missing requirements.
Thus, some requirements have been written at later design stages, as the ``Traceability (sourced from)'' column shows.
Also, the fourth column is to be completed in the subsystem design stage, when the system requirements will be allocated to one or more specific subsystems or components.


\section{Logical Decomposition}

The Logical Decomposition is an intermediate step between the Requirements Capture and the Design phases.
Its purpose is to understand the manner in which the requirements affect the way that the system functions, for the requirements loop; and to identify a feasible solution that functions in a way that meets the requirements, for the design loop, as shown in Figure \ref{fig:decomposition}

\input{./figures/decomposition.tex}

\subsection{Functional Architecture}

The logical decomposition performed during the functional analysis decomposes the top level requirements and allocates them down to the lowest desired levels.
The main outcome of the process is the Functional Architecture (Figure \ref{fig:functArch}), which helps establish relationships between requirements, and ultimately build a System Architecture.

\input{./figures/functArch.tex}

The main purpose is to create an association between the requirements and the functions that the system needs to be able to perform to meet them.
In the process, any discrepancy or missing items can (and should) be identified and corrected in an iterative manner.

\subsection{Functional Flow Block Diagram (FFBD)}

Once the functions of the system are defined, it is useful to dispose them so that the sequential use of each of them during the mission is shown.
To that end, the Functional Flow Block Diagram is used.
In the FFBD each function is represented by a block, and it is described in terms of inputs, outputs and interfaces.
In the case that a function is composed of several sub-functions, those will be represented hierarchicallyfrom the top level down to the most specific sub-function, maintaining the general flow.

The FFBD shows \emph{what} must happen, and provides an end-to-end path considering all the functionality of the system and the predefined use-case scenarios.
Parallel or alternate paths might be considered.

\input{./figures/ffbd.tex}

For the block diagrams depicted in Figures \ref{fig:ffbd0} to \ref{fig:ffbd6}, the symbology explained below is used:

\begin{itemize}
	\item 
		\tikz \draw[rounded corners=1mm,fill=teal!40,draw=black,semithick] (0,0) rectangle (6mm,3mm);
		\ represents an individual function or subfunction as defined in the Functional Architecture from Figure \ref{fig:functArch}.
	\item
		\tikz \draw[circle,fill=teal!70!black!70,draw=none,text=white,inner sep=1mm,node distance=2cm] (0,0) circle (2mm);
		\ represents a logical \textit{and} or \textit{or} gate for defining parallel or alternative paths, respectively.
	\item
		\tikz \draw[rectangle,rounded corners=1.5mm,fill=black!30,draw=black,thin] (0,0) rectangle (5mm,3mm);
		\ represents a reference block that specifies the origin or destination of a path from an external function of the system.
	\item
		\tikz \draw[rounded corners=1mm,draw=red!70!black!100,very thick,dashed] (0,0) rectangle (6mm,3mm);
		\ represents the boundaries of the functional description, be it the whole system or a subfunction of it.
	\item
		\tikz \draw[>=stealth,->,thick] (0,0) -- (5mm,0);
		\ indicates the secuential order that is to be followed from one function to another.
	\item
		\tikz \draw[>=stealth,->,thick,dashed] (0,0) -- (5mm,0);
		\ indicates an information flow between two functional blocks.
\end{itemize}


\subsection{Product Breakdown Structure (PBS)}

Once the functions are properly defined, they need to be allocated to the subsystems that will be in charge of accomplishing them.
To that end, the system is decomposed in its forming subsystems ensuring that all the functions can be achieved by the system.
The decomposition process is visually show via the Product Breakdown Structure, which is represented in Figure \ref{fig:pbs}

\input{./figures/pbs.tex}


\subsection{Functional-Physical matrix}

Finally, to couple the Requirements (Table \ref{tab:sysReqs}) with the Functional Architecture (Figure \ref{fig:functArch}) and with the Physical product (Figure \ref{fig:pbs}), a functional-physical matrix can be built.
This tool is very relevant since it exposes possible mismatches between the three steps, which would lead to requirements not being met or a product that cannot perform its intended functions.
Thus, by filling the matrix, the designer can go back to previous steps and adjust anything that is needed in order to avoid the exponential increase in cost that was mentioned at the beginning of the present Chapter.

For the matrix represented in Table \ref{tab:functPhys}, the requirements, functions and subsystems have been represented by their ID as defined in Table \ref{tab:sysReqs}, Figures \ref{fig:ffbd0} to \ref{fig:ffbd6} and Figure \ref{fig:pbs}, respectively.

\input{./tables/functPhys.tex}

As it can be seen for requirements 9.1, 10.1 and 11.1, they are requirements affecting the entire system, and thus cannot be associated to any function: they are requirements that impose hardware constrains only.

\subsection{Interfaces definition (N$^2$ diagram)}

For the correct integration of the system the definition of its interfaces is of utmost importance.
The N$^2$ diagram is commonly used for the development of those interfaces.

In the N$^2$ diagram, an N$\times$N matrix is built.
In the main diagonal all the systems or subsystems are placed, while the upper and lower triangles are reserved for the interfaces, which are clasified into inputs and outputs.
An input to any of the modules is represented through its vertical cells, while the output is place on the horizontal rows.
An example N$^2$ diagram can be found in Figure \ref{fig:N2example}

\input{./figures/N2example.tex}

For the Obstacle Collision Avoidance System N$^2$ diagram in Figure \ref{fig:N2}, the subsystems defined in the Product Breakdown Structure (Figure \ref{fig:pbs}) will be placed in the main diagonal.

\input{./figures/N2.tex}

In this Figure, the grey blocks represent external systems to the OCAS, while the blue ones are subsystems of the OCAS itself.
In a regular UAV, with no OCAS implemented, the pilot would only be executing the outermost green loop: only in direct control with the UAV.
Nonetheless, the OCAS provides an additional layer of safety that is functionally placed between the pilot and the UAV; but also adds another interface the pilot has to deal with: the GCS connector is the gate to the OCAS, which in turn connects via the Central Processing Unit to the rest of the system.
Hence, the ``green'' interfaces designate the human links with both the OCAS and the UAV.

Further, the ``yellow'' interfaces have been highlighted since they represent the traditional sensing and control problem.
Notice how the state of the UAV is transmitted from the UAV connector (that information is relayed by the UAV) and the sensors to the Processing Unit.
From the hardware-software interface downwards, the information is processed, the actuation calculated and ultimately the command is sent to the UAV. 

The ``blue'' interface has been created for information logging in response to Function 6.2 as defined in the FFBD.
It could also serve as a debugging channel during the implementation phase.

Finally, the ``black'' interfaces on the hardware side shows how the system has been designed meeting requirements 2.1 and 2.2, which stated that the OCAS shall be independent of the UAV and self-contained.
In this case, it has been decided that the OCAS would carry its own power source to provide energy to its components.
However, the power is not transmitted down to the UAV, which could be more weight efficient but would significantly modify the architecture of the original UAV, disregarding one of the main requirements from the motivation of the project.


\section{Component choice}

Only after properly defining the requirements, functions and subsystems of the Obstacle Collision Avoidance System can the components be chosen to ensure that the system is correctly designed to meet specifications.

\subsection{Sensors}

The most relevant sensing alternatives were already exposed in Section \ref{sec:sensing}.
In this Section, the most appropriate one for the project will be chosen according to the specifications.

Certainly, all the options considered have their advantages and disadvantages.
The purpose of the selection process is to evaluate all of those in a trade-off study which ensures that the final selection provides the best alternative to the system.
The steps to a successful trade-off study are:
\begin{enumerate}
	\item Definition of the problem
	\item Definition of the constraints
	\item Generation of alternative solutions
	\item Definition of the evaluation criteria
	\item Definition of the weight factors
	\item Fulfillment of the trade study
	\item Ranking of the solutions
\end{enumerate}

For the three first steps, those have already been done in Chapter \ref{chap:problem}, Section \ref{sec:sysReqs} and Section \ref{sec:sensing}, respectively.

For the rest of the steps, a tabular format will be used.

The evaluation criteria can be defined as the parameters considered important for the sensors to fulfill towards the correct achievement of their function, and will be presented on the first column of the table.

A weight factor will be given to each of the evaluation criteria according to their importance in the sensing and integration problems.
Those will be normalised ensuring that the sum of all of them add up to unity, and will be presented in the second column of the table.

The trade study will then be fulfilled by rating each of the alternatives on the evaluation criteria as previously defined.
The mark given, ranging from 0 to 1, represents the capability of the alternative on the corresponding evaluation criteria.

Finally, the ranking of the solutions is performed by combining (multiplying) the criteria's weight factors with the ratings given to the alternatives on each of those criteria, to ultimately sum all of those and obtain a single figure for every alternative considered.
The objectively most appropriate alternative is the one with highest rating after the trade-off study.

\input{./tables/tradeoff.tex}

